### 2020-03-20 CalNotes.tex
update delta=eps power

\subsection{Central limit theorem}

\corr{Text to be added with motivation, proof in Appendix \ref{ap:CLT}.
\begin{theorem}\label{thm:CLT} Central limit theorem statement:
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \sqrt{T}\left(\widehat A^\epl_k(T) - A\right) = \Lambda, \quad \text{in law,}
	\end{equation}
	where $\Lambda \sim \mathcal N(0, \Gamma)$ and
	\begin{equation}
		\Gamma = 2\sigma \E^{\rho^0}[V'(Z) \otimes V'(X)]^{-1}\E^{\rho^0}[V'(Z) \otimes V'(Z)]\E^{\rho^0}[V'(X) \otimes V'(Z)]^{-1}.
	\end{equation}
	This covariance may not be correct (see proof in the Appendix) but it is just the limit of the ``third term'' 
\end{theorem}
}

@@

\subsection{Central limit theorem}
\begin{figure}[t]
	\centering
	\includegraphics[]{Figures/CLT}
	\caption{Central limit theorem result. The histogram represents numerical results, the solid curve a Gaussian fit to the latter and the dashed curve the theoretical estimate given in Theorem \ref{thm:CLT}.}
	\label{fig:CLT}
\end{figure}

In this experiment we wish to confirm the validity of Theorem \ref{thm:CLT}. We consider the same test equation as for Section \ref{sec:Num_Param}, i.e., the quadratic potential $V(x) = x^2/2$ with fluctuating potential $p(y) = \sin(y)$, multiscale parameter $\epl = 0.05$ and diffusion coefficient $\sigma = 1$. The parameters of the filter are set to $\beta = 1$ and $\delta = 1$. We compute the estimator $\widehat A^\epl_k(T)$ with final time $T = 10^3$ on $2000$ realizations of the solution and estimate the quantity 
\begin{equation}	
	\Delta_A^\epl(T) \defeq \sqrt{T}\left(\widehat A^\epl_k(T) - A\right),	
\end{equation}
where $A$ is the drift coefficient of the homogenized equation. Results, depicted in Figure \ref{fig:CLT}, show that the distribution of $\Delta_A^\epl(T)$ indeed follows a zero-mean Gaussian law, \corr{whose covariance agrees with the theoretical results (?)}.

@@

\begin{lemma} \label{lem:zeta}
	For $\epl,\delta>0$ sufficiently small and $t>0$ sufficiently big, there exists a standard Brownian motion $B_t$ independent of the processes $X_t$ and $Z_t$ such that

@@

	X_t^\epl - Z_t^\epl = \sqrt{2\Sigma}\delta \int_0^t k(t-s) \dd B_s + R(\epl,\delta),

@@

	in law. The remainder $R(\epl,\delta)$ satisfies, for every $p > 0$ and $\iota \in (0,1/2)$, the estimate

@@

		\left( \E^{\rho_\epl} \left| R(\epl,\delta) \right|^p \right)^{1/p} \le C \left( \delta + \epl^\iota + \epl \delta^{1/2} \right),

@@

\begin{proof} Let $t > s$, $\Phi$ solution of \eqref{eq:CellProblem} and $Y_t^\epl = X_t^\epl / \epl$. Then, an application of the Itô formula yields \cite[Formula 5.8]{PaS07}
	\begin{equation}\label{eq:App_DiffXtXs}
	X_t^\epl - X_s^\epl &= -\int_s^t (\alpha \cdot V'(X_r^\epl))(1 + \Phi'(Y^\epl_r)) \dd r\\
	&\quad +\sqrt{2\sigma} \int_s^t (1 + \Phi'(Y_r^\epl)) \dd W_r - \epl(\Phi(Y^\epl_t)-\Phi(Y^\epl_s)).
	Recall that we have

@@

	Plugging \eqref{eq:App_DiffXtXs} into the equation above, we obtain

@@

	in law, where

@@

	Therefore, 

@@

	Now, Jensen's inequality with $\kappa(\d s)$ and the inequalities \eqref{eq:FirstTermPavStu} and \eqref{eq:FilterTminusS} yield for all integers $p \geq 1$ and a constant $C$ depending only on $p$

@@

	\E^{\phi^\epl}\abs{\widetilde R_1^\epl(t)}^{2p} &= (1-e^{-t/\delta})^{2p}\E^{\phi^\epl}\abs{\int_0^t R_1^\epl(t-s) \kappa(\d s)}^{2p}\\
	&\leq (1-e^{-t/\delta})^{2p} \int_0^t \E^{\phi^\epl} \abs{R_1^\epl(t-s)}^{2p} \kappa(\d s)\\
	&\leq C (1-e^{-t/\delta})^{2p-1} \int_0^t k(t-s)(\epl^{4p} + \epl^{2p}(t-s)^p+(t-s)^{3p}) \dd s\\
	&\leq C \left( \epl^{4p} + \epl^{2p} \delta^p + \delta^{3p} \right).

@@

	Therefore, for $t$ sufficiently big, we get

@@

		\left(\E^{\phi^\epl}\abs{I_1^\epl(t)}^{p}\right)^{1/p} \le& C \left( \delta + \epl^2 + \epl \delta^{1/2} + \delta^{3/2} \right) \\
		\le& C \left( \delta + \epl^2 + \epl \delta^{1/2} \right).

@@

	We now consider the second term. Let us remark that it holds \cite[Proposition 5.7, Section 6.2]{PaS07}
	\sqrt{2\sigma} \int_s^t \left(1+\Phi'(Y_s^\epl)\right) \dd W_r = \sqrt{2\Sigma}\left( B_t - B_s\right) + R_2^\epl(t-s),
	in law, where $\Sigma$ is the diffusion coefficient of the homogenized equation \eqref{eq:SDE_HOM}, $B_t$ is a standard Brownian motion \corr{independent of the process $X_t$ (?)} and 
	\left(\E^{\phi^\epl}\abs{R_2^\epl(t-s)}^p\right)^{1/p} \leq C\epl^\iota,
	for all $\iota \in (0, 1/2)$. Therefore, it holds in law
	\begin{equation} \label{eq:I2_decomposition}
	I_2^\epl(t) = \sqrt{2\Sigma} \int_0^t k(t-s) \left(B_t - B_s\right) \dd s + \widetilde R_2^\epl(t),
	\widetilde R_2^\epl(t) \defeq \sqrt{2\Sigma} \int_0^t k(t-s) R_2^\epl(t-s) \dd s.
	Applying Jensen's inequality with $\kappa(\d s)$ we obtain
	\left(\E^{\phi^\epl} \abs{\widetilde R_2^\epl(t)}^p\right)^{1/p} \leq C \epl^{\iota}.
	Let us remark that an application of the Itô formula yields
		\dd \left( k(t-s)B_s \right) = \frac1\delta k(t-s)B_s \dd s + k(t-s) \dd B_s,
	\begin{equation} \label{eq:Ito_kB}
		\int_0^t k(t-s)B_s \dd s = B_t - \delta \int_0^t k(t-s) \dd B_s.
	Substituting \eqref{eq:Ito_kB} into \eqref{eq:I2_decomposition} and recalling that
		\int_0^t k(t-s) \dd s = 1 - e^{-t/\delta},
	we obtain
		I_2^\epl(t) = \sqrt{2\Sigma} \delta \int_0^t k(t-s) \dd B_s - \sqrt{2\Sigma} e^{-t/\delta} B_t.
	Moreover note that
			\left(\E^{\phi^\epl} \abs{\sqrt{2\Sigma} e^{-t/\delta} B_t}^p\right)^{1/p} \le C e^{-t/\delta}.

	Let us now consider $I_3^\epl(t)$. Since $\Phi$ is bounded, we simply have
	\begin{equation}
	\abs{I_3^\epl(t)} \leq C \epl, 
	\end{equation}
	almost surely. Finally, due to \cite[Corollary 5.4]{PaS07}, we know that $X_t^\epl$ has bounded moments of all orders and therefore
	\begin{equation}
	\left(\E^{\phi^\epl} \abs{I_4^\epl(t)}^p\right)^{1/p} \leq C e^{-t/\delta},
	\end{equation}
	which concludes the proof.
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:mainTheorem_zeta}]
	Let us introduce the notation
	\begin{equation}
		\mathcal A^\epl(\delta) = \frac1\delta \widetilde{\mathcal M}_\epl^{-1} \E^{\rho^\epl}[(X^\epl - Z^\epl)^2 V''(Z^\epl)],
	\end{equation}
	where
	\begin{equation}
		\widetilde{\mathcal M}_\epl = \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ].
	\end{equation}
	Then following the proof of Theorem \ref{thm:mainTheorem} we only need to show that if $\delta = \epl^\zeta$ with $\zeta \in (0,1)$ we have
	\begin{equation}
		\lim_{\epl \to 0} \mathcal A^\epl(\delta) = A, \quad \text{in law}.
	\end{equation}
	Using Lemma \ref{lem:zeta} and geometric ergodicity (Lemma \ref{lem:ergodicity}), we have the following equality in law
	\begin{equation}
	\begin{aligned}
		\mathcal A^\epl(\delta) &= \widetilde{\mathcal M}_\epl^{-1} \frac1\delta \lim_{t \to \infty} \E [(X_t^\epl - Z_t^\epl)^2 V''(Z_t^\epl)] \\
		&= \widetilde{\mathcal M}_\epl^{-1} \frac1\delta \lim_{t \to \infty} \E \left[ \left( \sqrt{2\Sigma}\delta \int_0^t k(t-s) \dd B_s + R(\epl,\delta) \right)^2 V''(Z_t^\epl) \right] \\
		&\eqdef \widetilde{\mathcal M}_\epl^{-1} \lim_{t \to \infty} \left( J_1^\epl(t) + J_2^\epl(t) + J_3^\epl(t) \right),
	\end{aligned}
	\end{equation}
	where
	\begin{equation}
	\begin{aligned}
		J_1^\epl(t) &= 2\Sigma\delta \E^{\rho^\epl} \left[ \left( \int_0^t k(t-s) \dd B_s \right)^2 V''(Z_t^\epl) \right], \\
		J_2^\epl(t) &= 2\sqrt{2\Sigma} \E^{\rho^\epl} \left[ \left( \int_0^t k(t-s) \dd B_s \right) R(\epl,\delta) V''(Z_t^\epl) \right], \\
		J_3^\epl(t) &= \frac1\delta \E^{\rho^\epl} \left[ R(\epl,\delta)^2 V''(Z_t^\epl) \right].
	\end{aligned}
	\end{equation}
	Let us consider the three terms separately. First, since the Brownian motion $B_s$ is independent of the process $Z_t^\epl$, applying Itô isometry we get
	\begin{equation}
	J_1^\epl(t) = 2\Sigma\delta \left( \int_0^t k(t-s)^2 \dd s \right) \E [V''(Z_t^\epl)] = \Sigma \E^{\rho^\epl} [V''(Z_t^\epl)],
	\end{equation}
	and, by Lemma \ref{lem:distanceZandX} and weak convergence of the invariant measure, we obtain
	\begin{equation}
	\lim_{\epl \to 0} \lim_{t \to \infty} J_1^\epl(t) = \Sigma \lim_{\epl \to 0} \E^{\rho^\epl} [V''(Z^\epl)] = \Sigma \E^{\rho^0} [V''(X)].
	\end{equation}
	Moreover, due to \eqref{eq:remainder_zeta}, we have the following bounds for $J_2^\epl(t)$ and $J_3^\epl(t)$
	\begin{equation}
	\begin{aligned}
		\abs{J_2^\epl(t)} &\le C \left( \delta^{1/2} + \epl^\iota \delta^{-1/2} + \epl \right), \\
		\abs{J_3^\epl(t)} &\le C \left( \delta + \epl^{2\iota} \delta^{-1} + \epl^2 \right).
	\end{aligned}
	\end{equation}
	Furthermore, by Lemma \ref{lem:distanceMandTildeM} and weak convergence of the invariant measure, we have
	\begin{equation}
		\lim_{\epl \to 0} \widetilde{\mathcal M}_\epl^{-1} = \E^{\rho^0} [V'(X) \otimes V'(X) ] \eqdef \mathcal M_0.
	\end{equation}
	Therefore
	\begin{equation}
		\lim_{\epl \to 0} \mathcal A^\epl(\delta) = \Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ]^{-1} \E^{\rho^0} [V''(X)],
	\end{equation}
	and, finally, an integration by parts yields
	\begin{equation}
		\lim_{\epl \to 0} \mathcal A^\epl(\delta) = \Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ]^{-1} \frac1\Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ] A = A,
	\end{equation}
	which implies the desired result.
\end{proof}

\section{Proof of Theorem \ref{thm:CLT}}\label{ap:CLT}

Let us first introduce a technical Lemma.
\begin{lemma}\label{lem:CLT} Let $\mathcal L_\epl$ be the generator of the couple $(X^\epl, Z^\epl)^\top$, i.e.,
	\begin{equation}
		\mathcal L_\epl = -\left(\alpha \cdot V'(x) + \frac1\epl p'\left(\frac{x}{\epl}\right)\right)\partial_x + \frac1\delta (x - z) \partial_z  + \sigma \partial^2_{xx}.
	\end{equation} 
	Moreover, let $\rho^\epl$ be the density of the invariant measure of $(X^\epl, Z^\epl)^\top$ and $u^\epl\colon \R^2 \to \R^N$ be the solution of 
	\begin{equation}\label{eq:CLT_Poisson}
		-\mathcal L_\epl u^\epl = \chi^\epl -  \E^{\rho^\epl}[\chi^\epl(X^\epl, Z^\epl)],
	\end{equation}
	satisfying $\E^{\rho^\epl}[u^\epl(X^\epl, Z^\epl)] = 0$ for $\chi^\epl \colon \R^2 \to \R^N$. Then, it holds
	\begin{equation}\label{eq:CLT_PoissonIto}
		\frac{1}{T} \int_0^T  \chi^\epl(X_t^\epl, Z_t^\epl) \dd t = \E^{\rho^\epl}[\chi^\epl(X^\epl, Z^\epl)] - \frac{R^\epl(T)}{T} + \sqrt{2\sigma}\frac{S^\epl(T)}{T},
	\end{equation}
	where
	\begin{equation}\label{eq:CLT_PoissonRemainders}
		R^\epl(T) \defeq u^\epl(X^\epl_T, Z^\epl_T) - u^\epl(X^\epl_0, Z^\epl_0), \quad S^\epl(T) \defeq \int_0^T \partial_x u^\epl(X_t^\epl, Z_t^\epl) \dd W_t.
	\end{equation}
	Moreover, it holds
	\begin{equation}\label{eq:CLT_Equiv}
		2\sigma \E^{\rho^\epl}[\partial_x u^\epl(X^\epl, Z^\epl) \otimes \partial_x u^\epl(X^\epl, Z^\epl)] = \E^{\rho^\epl}
		\begin{aligned}[t]
		[&\chi^\epl(X^\epl, Z^\epl) \otimes u^\epl(X^\epl, Z^\epl) \\
		&+ u^\epl(X^\epl, Z^\epl) \otimes \chi^\epl(X^\epl, Z^\epl)].
		\end{aligned}
	\end{equation}	
\end{lemma}
\begin{proof} The proof of \eqref{eq:CLT_PoissonIto} and \eqref{eq:CLT_PoissonRemainders} is an application of the Itô formula (see e.g. \cite[Remark 6.17]{PaS08}). For \eqref{eq:CLT_Equiv}, it is possible to show that since $\mathcal L_\epl^*\rho^\epl = 0$ it holds 
	\begin{equation}
		\mathcal L_\epl^* (u^\epl \rho^\epl) = 2 \sigma \rho^\epl \partial_{xx}^2 u^\epl - \rho^\epl \mathcal L_\epl u^\epl + 2 \sigma \partial_x u^\epl \partial_x \rho^\epl.
	\end{equation}
	Therefore, an integration by parts yields 
	\begin{equation}
	\begin{aligned}
		\E^{\rho^\epl}[\mathcal L_\epl u^\epl(X^\epl, Z^\epl) \otimes u^\epl(X^\epl, Z^\epl)] &= 	\int_{\R}\int_{\R} u^\epl \otimes \mathcal L_\epl^*\left(u^\epl \rho^\epl\right) \dd x \dd z \\
		&= -\int_{\R}\int_{\R}u^\epl \otimes \mathcal L_\epl u^\epl \rho^\epl \dd x \dd z- 2\sigma \int_{\R}\int_{\R} \partial_x u^\epl \otimes \partial_x u^\epl \rho^\epl \dd x \dd z\\
		&= 
		\begin{aligned}[t]
		&-\E^{\rho^\epl}[u^\epl(X^\epl, Z^\epl) \otimes \mathcal L_\epl u^\epl(X^\epl, Z^\epl)] \\
		&- 2\sigma \E^{\rho^\epl}[\partial_x u^\epl(X^\epl, Z^\epl) \otimes \partial_x u^\epl(X^\epl, Z^\epl)].
		\end{aligned}
	\end{aligned}
	\end{equation}
	Finally, since $\E^{\rho^\epl}[u(X^\epl, Z^\epl)] = 0$ 
	\begin{equation}
	\begin{aligned}
		2\sigma \E^{\rho^\epl}[\partial_x u^\epl(X^\epl, Z^\epl) \otimes \partial_x u^\epl(X^\epl, Z^\epl)] &= -\E^{\rho^\epl}
		\begin{aligned}[t]
		[&\mathcal L_\epl u^\epl(X^\epl, Z^\epl) \otimes u^\epl(X^\epl, Z^\epl) \\
		& + u^\epl(X^\epl, Z^\epl) \otimes \mathcal L_\epl u^\epl(X^\epl, Z^\epl)]
		\end{aligned}
		\\
		&= \E^{\rho^\epl}
		\begin{aligned}[t]
		[&\chi^\epl(X^\epl, Z^\epl) \otimes u^\epl(X^\epl, Z^\epl) \\
		&+ u^\epl(X^\epl, Z^\epl) \otimes \chi^\epl(X^\epl, Z^\epl)],
		\end{aligned}
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

\begin{lemma} \label{lem:CLT_hom}
	Let $u^\epl$ be the solution of \eqref{eq:CLT_Poisson} with 
		\chi^\epl(x, z) = \frac 1 \epl p' \left ( \frac x \epl \right ) V'(z) - V'(z) \otimes V'(x) \mathcal M_\epl^{-1} \mathfrak p_\epl,
		\mathcal M_\epl \defeq \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ], \quad \mathfrak p_\epl \defeq \E^{\rho^\epl} \left [ \frac{1}{\epl} p' \left ( \frac{X^\epl}{\epl} \right ) V'(Z^\epl) \right ].
	Then, $u^\epl \to 0$ \corr{in which sense?} for $\epl \to 0$.
\begin{proof} We here present a formal proof based on asymptotic expansion with respect to $\epl$. Let us first remark that by definition
		\E^{\rho^\epl}[\chi^\epl(X^\epl, Z^\epl)] = 0,
	and therefore problem \eqref{eq:CLT_Poisson} reads
		-\mathcal L^\epl u^\epl = \chi^\epl.
	Let us now denote $y = x/\epl$ and write
		u^\epl(x, z) = u_0(x, y, z) + \epl u_1(x, y, z) + \epl^2 u_2(x, y, z) + \ldots,
	which implies that

@@

		\partial_x u^\epl &= \partial_x(u_0(x, y, z) + \epl u_1(x, y, z) + \epl^2 u_2(x, y, z) + \ldots) \\
		&\quad+ \frac1\epl\partial_y(u_0(x, y, z) + \epl u_1(x, y, z) + \epl^2 u_2(x, y, z) + \ldots).

@@

	Let us first remark that from the proof of Theorem \ref{thm:mainTheorem} we have that

@@

		\lim_{\epl \to 0} \mathcal M_\epl^{-1}\mathfrak p_\epl = A - \alpha.

@@

	Replacing $u^\epl$ in \eqref{eq:CLT_Poisson} and grouping the terms of order $\epl^0$, $\epl^{-1}$ and $\epl^{-2}$ we get the system
		L_0 u_0 &= 0, \label{eq:CLT_AsExp1}\\
		L_1 u_0 + L_0 u_1 &= - p'(y) V'(z), \label{eq:CLT_AsExp2}\\
		L_0 u_2 + L_1 u_1 + L_2 u_0 &= \left(V'(z) \otimes V'(x)\right) (A - \alpha),\label{eq:CLT_AsExp3}

@@

		L_0 &= -p'(y) \partial_y + \sigma \partial^2_{yy},\\
		L_1 &= -p'(y) \partial_x - \alpha \cdot V'(x) \partial_y + 2 \sigma \partial^2_{xy}, \\
		L_2 &= 	-\alpha \cdot V'(x) \partial_x - \frac1{\delta}(z - x)	\partial_z + \sigma \partial^2_{xx}.

@@

	Let us first remark that equation \eqref{eq:CLT_AsExp1} is satisfied for $u_0 = u_0(x, z)$ independent of $y$. In particular, the kernel of $L_0$ is made of constants and the kernel of $L_0^*$ is one-dimensional and $\mathrm{Ker}(L_0^*) = \mathrm{Span}\{\rho\}$ where
	\begin{equation}\label{eq:CLT_Invariant}
		\rho(y) = \frac1Z e^{-p(y)/\sigma}, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,

@@

	where $L$ is the period of $p$. Since $u_0$ is independent of $y$ equation \eqref{eq:CLT_AsExp2} reduces to 
		L_0 u_1 = p'(y)\left(\partial_x u_0 -V'(z)\right).
	Let us remark that the general solution $u_1$ can be written as
		u_1(x,y,z) = \Phi(y)\left(\partial_x u_0(x, z) - V'(z)\right),
	where $\Phi \colon \R \to \R$ satisfies
		-L_0 \Phi = -p'(y).
	Let us remark that this cell problem is the same as \eqref{eq:CellProblem}. We consider now equation \eqref{eq:CLT_AsExp3} and remark that we can rewrite it as

>>>>

@@

@@

\section{Estimates on the filtered solution}

In this section, we prove a series of results which quantify the distance between the multiscale solution $X_t^\epl$ and the filtered trajectory $Z_t^\epl$. First, we prove a Jensen-like inequality for the kernel of the filter.

\begin{lemma}\label{lem:FilterJensen} Let $\delta > 0$ and $k(r)$ be defined as
	\begin{equation}
		k(r) = \frac1\delta e^{-r/\delta}.
	\end{equation}
	Then, for any $t > 0$, $p \geq 0$ and any function $g\in \mathcal C^0(0, t)$ it holds 
	\begin{equation}
		\abs{\int_0^t k(t-s) g(s) \dd s}^p \leq (1 - e^{-t/\delta})^{p-1}\int_0^t k(t-s) \abs{g(s)}^p \dd s.
	\end{equation}
	Moreover, if $p \geq 1$ and since $0 < (1-e^{-t/\delta}) < 1$ for all $t > 0$, it holds
	\begin{equation}
		\abs{\int_0^t k(t-s) g(s) \dd s}^p \leq \int_0^t k(t-s) \abs{g(s)}^p \dd s.
	\end{equation}
\end{lemma}
\begin{proof} Let us first note that
	\begin{equation}
		\int_0^t k(t-s) \dd s = 1 - e^{-t/\delta}.
	\end{equation}
	Therefore, the measure $\kappa_t(\d s)$ on $(0, t)$ defined as
	\begin{equation}
		\kappa_t(\d s) \defeq \frac{k(t-s)}{1 - e^{-t/\delta}} \dd s,
	\end{equation}
	is a probability measure. An application of Jensen's inequality therefore yields
	\begin{equation}
	\begin{aligned}
		\abs{\int_0^t k(t-s) g(s) \dd s}^p &\leq (1 - e^{-t/\delta})^p \int_0^t \abs{g(s)}^p \kappa_t(\d s) \\
		&= (1 - e^{-t/\delta})^{p-1} \int_0^t k(t-s) \abs{g(s)}^p \dd s,
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

The following lemma characterizes the action of the filter when it is applied to polynomials.
\begin{lemma}\label{lem:FilterPoly} With the notation of Lemma \ref{lem:FilterJensen}, it holds for all $p \geq 0$
	\begin{equation}
		\int_0^t k(t-s)(t-s)^p \dd s \leq C \delta^p,
	\end{equation}
	where $C > 0$ is a positive constant independent of $\delta$.
\end{lemma}
\begin{proof} The change of variable $u = (t-s)/\delta$ yields
	\begin{equation}
		\int_0^t k(t-s)(t-s)^p \dd s = \delta^p \int_0^{t/\delta} u^p e^{-u} \dd u = \delta^p \gamma\left(p+1, \frac{t}{\delta}\right),
	\end{equation}
	where $\gamma$ is the lower incomplete Gamma function, which is bounded by the complete Gamma function $\Gamma(p+1)$ independently of the second argument.
\end{proof}
We can now prove a rough estimate of the difference between the trajectories $X_t^\epl$ and $Z_t^\epl$ when they are at stationarity.

\begin{lemma}\label{lem:distanceZandX} Under Assumption \ref{as:regularity}, let the couple $(X^\epl, Z^\epl)^\top$ be distributed as its invariant measure $\mu^\epl$. Then, if $\delta \leq 1$, it holds for any $p \geq 1$
	\begin{equation}
		\left(\E^{\rho^\epl} \abs{X^\epl - Z^\epl}^p\right)^{1/p} \leq C\left(\epl + \delta^{1/2}\right),
	\end{equation}
	for a constant $C > 0$ independent of $\epl$ and $\delta$.
\end{lemma}
\begin{proof} Let $X_t^\epl$ be at stationarity with respect to its invariant measure, which we recall having density denoted as $\phi^\epl$. Let $Z_t^\epl$ be the corresponding filtered process. By definition of $Z^\epl_t$ we have
	\begin{equation}
	\begin{aligned}
		X_t^\epl - Z_t^\epl &= X_t^\epl - \int_0^t k(t-s) X_s^\epl \dd s\\
		&= \int_0^t k(t - s) \left(X_t^\epl - X_s^\epl\right) \dd s + e^{-t/\delta} X^\epl_t.
	\end{aligned}
	\end{equation}
	Therefore, Lemma \ref{lem:FilterJensen} yields for a constant $C > 0$ depending only on $p$
	\begin{equation}
		\E^{\phi^\epl}\abs{X_t^\epl - Z_t^\epl}^p \leq C\int_0^t \E^{\phi^\epl} \abs{X_t^\epl - X_s^\epl}^p k(t-s) \dd s + Ce^{-pt/\delta} \E^{\phi^\epl} \abs{X^\epl_t}^p.
	\end{equation}
	Moreover, the estimate
	\begin{equation}
		\E^{\phi^\epl}\abs{X_t^\epl - X_s^\epl}^p \leq C\left(\epl^p + (t-s)^p + (t-s)^{p/2}\right),
	\end{equation}
	holds for all $t > s$ due to \cite[Lemma 6.1]{PaS07}. Finally, since the moments of $X_t^\epl$ are bounded \cite[Corollary 5.4]{PaS07} and due to Lemma \ref{lem:FilterPoly} there exists a constant $C > 0$ depending only on $p$ such that
	\begin{equation}
		\E^{\phi^\epl}\abs{X_t^\epl - Z_t^\epl}^p \leq C\left(\epl^p + \delta^{p/2} + e^{-pt/\delta}\right).
	\end{equation}
	Let us now remark that this result holds for $X_t^\epl$ being at stationarity and for $Z_t^\epl$ be its filtered process, and not for a couple $(X^\epl, Z^\epl)^\top \sim \mu^\epl$. In order to conclude, we remark that due to Lemma \ref{lem:ergodicity} we have for all $t \geq 0$ and for a constant $C$ depending on the initial condition and $p$  
	\begin{equation}
	\E^{\rho^\epl} \abs{X^\epl - Z^\epl}^p \leq \E^{\phi^\epl}\abs{X_t^\epl - Z_t^\epl}^p + Ce^{-\lambda t},
	\end{equation}
	which, for $t$ sufficiently big, yields the desired result.
\end{proof}

It is relevant to study the quantities
\begin{equation}\label{eq:CalM}
\mathcal M_\epl \defeq \E^{\rho^\epl}[V'(X^\epl) \otimes V'(X^\epl)] \quad \text{and} \quad \widetilde{\mathcal M}_\epl \defeq \E^{\rho^\epl}[V'(Z^\epl) \otimes V'(X^\epl)],
\end{equation}
which are by the ergodic theorem the time limit of the matrices appearing in the estimators $\widehat A^\epl(T)$ and $\widehat A^\epl_k(T)$, respectively. In particular, the distance between the matrices $\mathcal M_\epl$ and $\widetilde{\mathcal M}_\epl$ can be bounded, which is summarized in the following lemma.

\begin{lemma}\label{lem:distanceMandTildeM} Let the assumptions of Lemma \ref{lem:distanceZandX} hold. Then the matrices $\mathcal M$ and $\widetilde{\mathcal M}_\epl$ defined in \eqref{eq:CalM} satisfy
	\begin{equation}
	\norm{\mathcal M_\epl - \widetilde{\mathcal M}_\epl}_2 \leq C\left(\epl + \delta^{1/2}\right),
	\end{equation}
	for a constant $C > 0$ independent of $\epl$ and $\delta$.
\end{lemma}
\begin{proof} Applying Jensen's and Cauchy--Schwarz inequalities we have
	\begin{equation}
	\begin{aligned}
	\norm{\mathcal M_\epl - \widetilde{\mathcal M}_\epl}_2 &\leq \E^{\rho^\epl} \norm{\left(V'(Z^\epl) - V'(X^\epl)\right) \otimes V'(X^\epl)}_2 \\
	&\leq \left(\E^{\rho^\epl}\norm{V'(Z^\epl) - V'(X^\epl)}_2^2\right)^{1/2} \left(\E^{\rho^\epl} \norm{V'(X^\epl)}_2^{2}\right)^{1/2}.
	\end{aligned}
	\end{equation}
	The Lipschitz condition on $V'$ together with the boundedness of the moments of $X^\epl$ and Lemma \ref{lem:distanceZandX} yield for a constant $C > 0$
	\begin{equation}
	\begin{aligned}
	\norm{\mathcal M_\epl - \widetilde{\mathcal M}_\epl}_2 &\leq C\left(\E^{\rho^\epl}\abs{Z^\epl - X^\epl}^{2}\right)^{1/2} \leq C \left(\epl + \delta^{1/2} \right),
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

The final result presented in this section is a sharper quantification of the difference between $X^\epl_t$ and $Z^\epl_t$ than Lemma \ref{lem:distanceZandX}. In particular, denoting $Y_t^\epl \defeq X_t^\epl/\epl$, we will make use of the decomposition
\cite[Formula 5.8]{PaS07}
\begin{equation}\label{eq:App_DiffXtXs}
\begin{aligned}
	X_t^\epl - X_s^\epl &= -\int_s^t (\alpha \cdot V'(X_r^\epl))(1 + \Phi'(Y^\epl_r)) \dd r\\
	&\quad +\sqrt{2\sigma} \int_s^t (1 + \Phi'(Y_r^\epl)) \dd W_r - \epl(\Phi(Y^\epl_t)-\Phi(Y^\epl_s)),
\end{aligned}
\end{equation}
which is obtained applying the Itô formula to $\Phi$, the solution of the cell problem \eqref{eq:CellProblem}.

\begin{lemma} \label{lem:zeta} Let Assumption \ref{as:regularity} hold and $\epl,\delta>0$ be sufficiently small. Then, it holds for every $t > 0$ sufficiently big 

@@

		X_t^\epl - Z_t^\epl = \delta \beta^\epl_t + R(\epl,\delta),

@@

	where the stochastic process $\beta_t^\epl$ is defined as
	\begin{equation}\label{eq:defBeta}
		\beta_t^\epl \defeq \sqrt{2\sigma} \int_0^t k(t-s)(1 + \Phi'(Y_s^\epl)) \dd W_s,
	where $\Phi$ is the solution of the cell problem \eqref{eq:CellProblem}, $W_s$ is the Brownian motion appearing in \eqref{eq:SDE_MS} and $Y_t^\epl = X_t^\epl / \epl$. Moreover, the remainder $R(\epl,\delta)$ satisfies for every $p > 0$ the estimate

@@

	\left( \E^{\rho_\epl} \left| R(\epl,\delta) \right|^p \right)^{1/p} \le C \left( \delta + \epl^2 + \epl \delta^{1/2} \right),

@@

\begin{proof} Recall that by definition of $Z_t^\epl$ we have

@@

	Plugging the decomposition \eqref{eq:App_DiffXtXs} into the equation above, we obtain

@@

	in law, where the remainder $R_1^\epl$ satisfies 

@@

	Therefore, it holds 

@@

	Now, Lemma \ref{lem:FilterJensen}, the inequality \eqref{eq:FirstTermPavStu} and Lemma \ref{lem:FilterPoly} yield for all $p \geq 1$

@@

		\E^{\phi^\epl}\abs{\widetilde R_1^\epl(t)}^p &\leq C \int_0^t k(t-s)\E^{\phi^\epl} \abs{R_1^\epl(t-s)}^p \dd s\\
		&\leq C \int_0^t k(t-s)(\epl^{2p} + \epl^p(t-s)^{p/2}+(t-s)^{3p/2}) \dd s\\
		&\leq C \left( \epl^{2p} + \epl^p \delta^{p/2} + \delta^{3p/2} \right),

@@

	where $C$ is a positive constant independent of $\epl$ and $\delta$. Therefore, for $t$ sufficiently big and $\delta$ sufficiently small, we get
		\left(\E^{\phi^\epl}\abs{I_1^\epl(t)}^{p}\right)^{1/p} \leq C \left( \delta + \epl^2 + \epl \delta^{1/2} \right).
	We now consider the second term. Let us introduce the notation  
		Q_t \defeq \int_0^t \left(1 + \Phi'(Y_r^\epl)\right) \dd W_r,
	and therefore rewrite
		I_2^\epl(t) = \sqrt{2\sigma} Q_t \int_0^t k(t-s)(Q_t - Q_s) \dd s.
	An application of the Itô formula to $u(s, Q_s)$ where $u(s, x) = k(t-s)x$ yields

@@

		I_2^\epl(t) &= \sqrt{2\sigma} \left(Q_t\int_0^t k(t-s) \dd s - Q_t + \delta \int_0^t k(t-s)\left(1 + \Phi'(Y_s^\epl)\right) \dd W_s\right)\\
		  			&= \delta \beta_t^\epl - \sqrt{2\sigma} e^{-t/\delta} Q_t \eqdef \delta \beta_t^\epl - R_2^\epl(t).

@@

	where $\beta_t^\epl$ is defined in \eqref{eq:defBeta}. For the remainder $R_2^\epl(t)$, let us remark that for all $p \geq 1$ it holds

@@

		\left(\E\abs{Q_t}^p\right)^2 \leq \E\abs{Q_t}^{2p} \leq Ct^{p-1}\int_0^t \E \abs{1 + \Phi'(Y_r^\epl)}^{2p} \dd r \leq C t

@@

	where we applied Jensen's inequality, an estimate for the moments of stochastic integrals \cite[Formula (3.25), p. 163]{KaS91} and the boundedness of $\Phi$. Therefore, for $t$ large enough, we have

@@

		\E\abs{R_2^\epl(t)}^p \leq C \sqrt{t}e^{-pt/\delta}.

@@

	Let us now consider $I_3^\epl(t)$. Since $\Phi$ is bounded, we simply have

@@

	\abs{I_3^\epl(t)} \leq C \epl, 

@@

	almost surely. Finally, due to \cite[Corollary 5.4]{PaS07}, we know that $X_t^\epl$ has bounded moments of all orders and therefore
	\left(\E^{\phi^\epl} \abs{I_4^\epl(t)}^p\right)^{1/p} \leq C e^{-t/\delta},

@@

	which concludes the proof.
